# Using BigQuery

## Data

We will use the NYC Bicycle Hire data we looked at in lecture. The database contains all Citi Bike trips (NYC's bike-sharing service) since Citi Bike launched in September 2013. 

```{r}
library("DBI")
library("bigrquery")
```


## Queries

```{r}
project_name <- "bigquery-public-data"

billing_name <- "third-pen-333710"

dataset_name <- "new_york_citibike"
```


1. First, connect to the database and see the first 5 entries.

```{r}
db <- dbConnect(bigrquery::bigquery(),
                project = project_name,
                billing = billing_name,
                dataset = dataset_name)
dbGetQuery(db, 
           "SELECT * FROM citibike_trips LIMIT 5")
```

2. What is the average trip duration based on the age of the riders?

```{r}
dbListFields(db, "citibike_trips")
```


```{r}
dbGetQuery(db,
           "SELECT birth_year, AVG(tripduration) / 60 AS avg_trip_duration
           FROM citibike_trips
           GROUP BY birth_year
           ORDER BY avg_trip_duration")

```

3. What is the average distance of a CitiBike trip based on the age of the riders?

```{r}
dbGetQuery(db,
          "SELECT birth_year, AVG( 
          (ABS(end_station_longitude - start_station_longitude) + 
           ABS(end_station_longitude -end_station_longitude) )
           * 111) AS trip_distance
           FROM citibike_trips
           GROUP BY birth_year
           ORDER BY trip_distance DESC")

```

4. What is the average speed of a CitiBike trip?

```{r}
dbGetQuery(db,
          "SELECT AVG(trip_distance) / AVG(hours) AS speed_km_h
          FROM(SELECT 
          (ABS(end_station_longitude - start_station_longitude) + 
           ABS(end_station_longitude -end_station_longitude) )
           * 111 AS trip_distance, 
           tripduration / 3600.0 AS hours
           FROM citibike_trips)")


```

5. What is the average speed based on age?

```{r}
dbGetQuery(db,
          "SELECT birth_year,
          AVG(trip_distance) / AVG(hours) AS speed_km_h
          FROM(SELECT birth_year,
          (ABS(end_station_longitude - start_station_longitude) + 
           ABS(end_station_longitude -end_station_longitude) )
           * 111 AS trip_distance, 
           tripduration / 3600.0 AS hours
           FROM citibike_trips)
          GROUP BY birth_year
          ORDER BY speed_km_h DESC")

```

