# Subqueries in SQL


## Facebook database

Use the SQLite database with Facebook posts data we created and used in week 10 to complete the questions below.

## Subqueries

In SQL, you can run a nested query where you use a result from the first query as a condition for the second query. Here, we will find the top 3 most frequent posters in congress and then get all messages from them. 

```{r}
# load required libraries and connect to facebook-db.sqlite
library(DBI)
library(RSQLite)
library(microbenchmark)
db <- dbConnect(RSQLite::SQLite(), "data/facebook-db-mini.sqlite")
```

1. What are the screen names of the top 3 posters?
```{r}
# It only shows the screen names of the politicians who posted more or equal than the FIRST of the three. WHen using HAVING >= How can I solve this?
# When using IN we retrieve all politicians who have these post counts. We only want the three that appear.
dbGetQuery(db,
           "SELECT screen_name
           FROM posts
           GROUP BY screen_name
           HAVING COUNT(*) IN (SELECT COUNT(*)
           FROM posts
           GROUP BY screen_name
           ORDER BY COUNT(*) DESC
           LIMIT 3)")

# Here is how I would do it without a subquery.
dbGetQuery(db,
           "SELECT screen_name
           FROM posts
           GROUP BY screen_name
           ORDER BY COUNT(*) DESC
           LIMIT 3")
```

2. (Without subquery) how can you get the posts from these congress members?
```{r}
dbGetQuery(db,
           "SELECT *
           FROM posts
           WHERE screen_name IN ('CongressmanNadler', 'SenatorHeidiHeitkamp', 'RepMikeThompson')")

```

3. Now, instead of hardcoding the screen_names in the conditional statement, use the query results in #1 as the subquery. 
```{r}
# With subquery
dbGetQuery(db,
           "SELECT *
           FROM posts
           WHERE screen_name IN (SELECT screen_name 
           FROM posts
           GROUP BY screen_name
           ORDER BY COUNT(*) DESC
           LIMIT 3)
           ")
```

4. Using a subquery, get the average number of likes per post for all women Democrats.
```{r}
dbGetQuery(db,
           "SELECT AVG(likes_count)
           FROM posts
           WHERE screen_name IN (
           SELECT screen_name
           FROM congress
           WHERE gender = 'F' AND party = 'Democrat'
           )"
           )
```

5. Now use a join to do the same as in #4.
```{r}
dbGetQuery(db,
           "SELECT AVG(posts.likes_count)
           FROM congress JOIN posts
           ON congress.screen_name = posts.screen_name
           WHERE congress.gender = 'F' AND congress.party = 'Democrat'")
```

6. Which solution is faster: the one in #4 or #5?

```{r}
microbenchmark(dbGetQuery(db,
           "SELECT AVG(likes_count)
           FROM posts
           WHERE screen_name IN (
           SELECT screen_name
           FROM congress
           WHERE gender = 'F' AND party = 'Democrat'
           )"), times = 100)
microbenchmark(dbGetQuery(db,
           "SELECT AVG(posts.likes_count)
           FROM congress JOIN posts
           ON congress.screen_name = posts.screen_name
           WHERE congress.gender = 'F' AND congress.party = 'Democrat'"),
           times = 100)

# Median of 6 milliseconds compared with a median of 18 milliseconds. Using subqueries is three times faster than using joins.
```
