---
title: "ggplot2 Basics: Grammar of graphics, geom types"
author: "Pablo Barberá and Friedrich Geiecke"
date: "15/11/2021"
output: github_document
---

## Introduction

In the examples below, we will be working with a random sample of 10,000 public Facebook posts by members of the U.S. Congress. 

For a review of the log scale, see for example this [video](https://www.youtube.com/watch?v=sBhEi4L91Sg).

Loading packages:

```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("lubridate")
library("scales")
```

Reading in the data and some initial processing:

```{r}
df <- read.csv("data/fb-congress-data.csv", stringsAsFactors = FALSE) %>% as_tibble()

# Transform date column to datetime
df$date <- as_date(df$date)

# Dropping some very uncommon post types
nrow(df)
df <- df %>%
  filter(!post_type %in% c("music", "note"))
nrow(df)
```


## Time series

After creating a base layer, `geom_line()` can be used for line plots such as time series. Plotting daily posts over time:

```{r}
counts <- df %>%
  group_by(date) %>% 
  summarise(posts = n())

# Base layer
p <- ggplot(counts, aes(x = date, y = posts))

# Line plot of the posts per day
p + geom_line()
```


Two separate time series by party, now also aggregated to monthly:

```{r}
# Obtain a new data frame with monthly counts of posts per party
counts <- df %>%
  filter(party != "Independent") %>% 
  # ceiling_date function aggregates the days into months, with the next unit that encompasses it as a representation.
  # for instance, here the dates that are represented as the next month (because they are in the interval of month m and m + 1, so they will be shown as month m + 1).
  group_by(month = ceiling_date(date, "month"), party) %>% 
  summarise(posts = n())

p <- ggplot(counts, aes(x = month, 
                        y = posts
                        ))
p + geom_line(aes(color = party)) +
  scale_color_manual(values=c(Republican = "blue", Democrat = "red"))

# Here there is no need to use the group = party argument.
# Here the order of the values is assigned depending on which value appears first in the dataset. Since Democrat appears first, it is the first color value. You can also change this specification. 
```

**Explanation of the scale_x_log10. What this does is: It takes the logarithm of the x-values, creates the figure with these logarithms, and then uses the previous linear axis. It is as if we transformed the numbers but made a reference to the previous values (before transformation). Some numbers fall out of the figure because of the xlims, but we can change this. THis is useful when we want to generate transformations without eliminating readability of our figure.**
**You can notice that it is a logarithmic scale because the axis shows the same distance between 10-100 and between 100-1000.**
**You can read the y values in the y-axis as exponents of log y. (The values are logs of y, whereas the values displayed in the axes are only y).**

## Univariate analysis for a single continuous variable

```{r}
# Base layer
p <- ggplot(df, aes(x=likes_count))

# Histogram
p + geom_histogram()

# Smoothed density estimate
p + geom_density() + scale_x_continuous("likes count", labels = comma)

# The same log scale (note: the labels = comma option prevents scientific
# notation of numbers as 1e+00, 1e+01 and uses the `scales` package)
# Log allows really skewed distributions to appear normal (log-normal).
p + geom_histogram() + scale_x_log10("likes count", labels = comma)
p + geom_density() + scale_x_log10("likes count", labels = comma)

# Why does this line of code drop some observations?
# Because these are the posts with a like count of zero. Hence, for these the logarithm is an undetermined value, so our geoms directly remove them from the plotting dataset. 
```

```{r}
p1 <- ggplot(mpg, aes(displ, hwy)) +
    geom_point()
 p1
 p1 + scale_y_log10()
 p1 + scale_x_log10()
 p1 + scale_y_log10() + scale_x_log10()
 ggplot(mpg, aes(x = log(displ, base = 10), y = log(hwy, base = 10))) + geom_point()
```



## Univariate analysis for a single categorical variable

```{r}
p <- ggplot(df, aes(x=post_type)) + xlab("post type")

# Bar chart
p + geom_bar() 
## number of posts by type. This is already performing a stats_count() function
# under the hood

# Bar chart (horizontal)
p + geom_bar() + coord_flip()
```


## Bivariate analysis for two continuous variables

```{r}
# Base layer
p <- ggplot(df, aes(x = likes_count, y = comments_count)) + xlab("Likes count") +
  ylab("Comments count")

# Scatter plot: Relationship between number of likes and number of comments
p + geom_point()  

# With smoothed conditional means
p + geom_point() + stat_smooth(na.rm = TRUE) 

# With restricted axes
p + geom_point() + xlim(0, 25000) + ylim(0, 2500)

# Particularities of integer variables in scatter plots
p + geom_point() + xlim(0, 10) + ylim(0, 10)
  
# With log scales
p + geom_point() + scale_x_log10(labels = comma) + scale_y_log10(labels = comma) 

p + geom_point() + scale_x_log10(labels = comma) + scale_y_log10(labels = comma) + 
    stat_smooth()
```

When we apply the logplot, there is an unusual shape at the beginning of the plot (because of the combination of the integer counts and the logscale). This happens because of: 1) The grid shape (which is shown at the beginning), and 2) The smaller density of points at the beginning of the range of values (there are less posts with likes count between [0-10] than [10-1000]), and this causes the grid shape to be hidden for the middle of the graph.

## Bivariate analysis for one continuous variable and one categorical variable

```{r}
# Number of likes by type of post as a box blot
p <- ggplot(df, aes(x = post_type, y = likes_count)) + xlab("Post type") +
  ylab("Likes count")
p + geom_boxplot()

# Box plot and violin plot with log scale
# When you cannot see anything, you need a log transformation. 
p + geom_boxplot() + scale_y_log10(labels = comma)
p + geom_violin() + scale_y_log10(labels = comma)

# Density plot for log like distributions for different parties
p <- ggplot(df, aes(x = likes_count))
p + geom_density(aes(color = party)) + scale_x_log10("likes count", labels = comma)
```

To calculate the density function, you create a histogram and fit the density function to it. 

## Bivariate analysis for two categorical variables

```{r}
counts <- df %>%
  filter(party != "Independent") %>% 
  group_by(post_type, party) %>% 
  summarise(posts = n())

p <- ggplot(counts, aes(x = party, y = post_type)) + ylab("post type")
p + geom_tile(aes(fill = posts))

```
Republicans post more photo types. 

## Multivariate analysis for three continuous variables

```{r}
p <- ggplot(df, aes(x = likes_count, y = comments_count, color = log(angry_count))) +
  xlab("Likes count") + ylab("Comments count")

p + geom_point()

p + geom_point() + scale_y_log10(labels = comma) + scale_x_log10(labels = comma) + 
 stat_smooth(method = "lm")
```

Angry reactions are more popular.

## Multivariate analysis for two continuous variables and one categorical variable

```{r}

# Grid of plots: 2x4, by post type
p <- ggplot(df, aes(x = likes_count, y = comments_count)) + xlab("Likes count") +
  ylab("Comments count")
p + geom_point() + scale_x_log10(labels = comma) + scale_y_log10(labels = comma) + 
        facet_wrap(~post_type, nrow = 2) 


# geom_text() allows to use party names instead of points
p <- ggplot(df[df$likes_count>10000, ], 
            aes(x = likes_count, y = comments_count, label = party)) +
  xlab("Likes count") +
  ylab("Comments count")
p + geom_text() + scale_x_log10(labels = comma) + scale_y_log10(labels = comma)
```

Other examples:

```{r}
## Scatter plot with dots colored by type of post
p <- ggplot(df[df$likes_count>5000, ], 
            aes(x = likes_count, y = comments_count)) +
  scale_x_log10("Likes count", labels = comma) +
  scale_y_log10("Comments count", labels = comma)
p + geom_point(aes(color = post_type))

## Same for point shape
p + geom_point(aes(shape = post_type))

## Combining both (now different shapes also have different colors)
# You can combine both!
p + geom_point(aes(shape = post_type, color = post_type))
```

## Dealing with cases where a lot of points are in some areas

Jittering points can avoid "overplotting", however, can also easily be misleading:

```{r}
p <- ggplot(df, aes(x = party, y = comments_count)) + ylab("comments count")
p + geom_point()

# vs
# position_jitter shows you how to spread the jittered values (the margins on width and height).
p + geom_jitter(position = position_jitter(width = .1, height=.1))
```

```{r}
# Baseline
p <- ggplot(df, aes(x = likes_count, y = comments_count)) +
  scale_x_log10("Likes count", labels = comma) +
  scale_y_log10("Comments count", labels = comma) 
p + geom_point()

## Jittering points (yet, really changes the plot's appearance here and can be misleading because it looks like the variables are of double type).
p + geom_jitter(position = position_jitter(width = .5, height =.5))

## Transparency (lower alpha)
p + geom_jitter(position = position_jitter(width = .5, height = .5), alpha = 1/25)

## Hexbin (if error: run install.packages("hexbin"))
# This bins values
p <- ggplot(df[df$likes_count>0 & df$comments_count>0,], 
        aes(x=likes_count, y = comments_count))
p + geom_hex() + scale_x_log10() + scale_y_log10() # scale_fill_continuous(trans="log10")

# Generally, plotting binned means (either over deciles or over fixed (log) intervals like here)
# can help with plots that contain a lot of points/observations.
# Furthermore, geom_rug() is another option to indicate where most mass is
# Geom_rug is in the sides, and stat_summary_bin shows the mean of each bin (for twenty bins across the whole variable domain),
p <- ggplot(df, aes(x = likes_count, y = comments_count)) +
  scale_x_log10("Likes count", labels = comma) +
  scale_y_log10("Comments count", labels = comma) 
p + geom_point() + geom_rug(color = "grey", alpha = 0.6) +
   stat_summary_bin(fun = 'mean', bins = 20, color = 'green', size = 2, geom = 'point')
# You can do this with equally distanced bins (as in this case), or 
# by percentiles.

```
