# Practicing working with multiple SQL tables

Practice writing using **JOIN** in SQL queries to answer the questions below. Use the database we created in lecture.

1. How many posts were published each month of the year?

```{r}
library(DBI)
library(RSQLite)
db <- dbConnect(RSQLite::SQLite(), "data/facebook-db.sqlite")

dbGetQuery(db, 
           "SELECT SUBSTR(date, 6, 2) AS month, COUNT(*) AS n_posts
           FROM posts 
           GROUP BY month")

```


2. What is the average number of likes for members of the House and the Senate? Which group is more popular on average?

```{r}
dbGetQuery(db,
           "SELECT congress.type, AVG(posts.likes_count) 
           FROM congress JOIN posts
           ON congress.screen_name = posts.screen_name
           GROUP BY congress.type")
```

3. What is the average number of likes that posts receive based on the time of the day in which they were published?

```{r}
dbGetQuery(db,
           "SELECT SUBSTR(datetime, 12, 2) AS hour, AVG(likes_count)
           FROM posts 
           GROUP BY hour")

```

4. Do women receive more comments than men on average?

```{r}
dbGetQuery(db,
           "SELECT congress.gender, AVG(posts.comments_count)
           FROM congress JOIN posts
           ON congress.screen_name = posts.screen_name
           GROUP BY congress.gender")
```

5. For each type of reaction (haha, love, angry...) compute the average count by party. Are there any systematic differences?

```{r}
dbGetQuery(db,
           "SELECT congress.party, AVG(posts.likes_count), AVG(posts.comments_count), AVG(posts.shares_count),
           AVG(posts.love_count), AVG(posts.haha_count), AVG(posts.wow_count), AVG(posts.sad_count), AVG(posts.angry_count)
           FROM congress JOIN posts
           ON congress.screen_name = posts.screen_name
           GROUP BY congress.party")
```

6. Compute the total number of "sad" reactions for each domain shared by Republicans. Display the top 10 domains that receive the most "sad" reactions. Then repeat the same for Democrats. What do you find?

```{r}
dbGetQuery(db,
           "SELECT posts.domain, SUM(posts.sad_count) AS n_sad_reactions
           FROM congress JOIN posts
           ON congress.screen_name = posts.screen_name
           WHERE congress.party = 'Republican' AND posts.domain NOT NULL
           GROUP BY posts.domain
           ORDER BY n_sad_reactions DESC
           LIMIT 10")

dbGetQuery(db,
           "SELECT posts.domain, SUM(posts.sad_count) AS n_sad_reactions
           FROM congress JOIN posts
           ON congress.screen_name = posts.screen_name
           WHERE congress.party = 'Democrat' AND posts.domain NOT NULL
           GROUP BY posts.domain
           ORDER BY n_sad_reactions DESC
           LIMIT 10")
```






